---
title: "Text Mining: Sentiment analysis"
author: "Amalia Martínez"
date: "14 October 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(kableExtra)
```

## 

Code available at 
<www.github.com/amarseg/decoded_presentation/>

## What is sentiment analysis?

"The process of computationally identifying and categorizing opinions expressed in a piece of text, especially in order to determine whether the writer's attitude towards a particular topic, product, etc. is positive, negative, or neutral."

What can it be used for?

- Monitor social media
- Extract information from online reviews of the product
- Creation of a recommendation system in an online marketplace

## Loading libraries and datasets

```{r, eval = F, echo = T,warning=F, message=F}
install.packages('tidyverse')
install.packages('tidytext')
install.packages('janeaustenr')
install.packages('wordcloud')
```


```{r, echo = TRUE,warning=F, message=F}
library(tidyverse)
library(tidytext)
library(janeaustenr)
```

## Jane Austen dataset
```{r}
head(austen_books(), n = 15)
```

## Transform data into a format amenable for analysis

```{r, echo = TRUE,warning=F, message=F}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter",
                                           ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

## Transform data into a format amenable for analysis
```{r}
tidy_books 
```
## Affinn sentiment datasets
From Finn Årup Nielsen
```{r, echo = T}
get_sentiments("afinn")
```

## NRC Sentiment datasets
from Saif Mohammad and Peter Turney
```{r, echo = T}
get_sentiments("nrc")
```

## BingSentiment datasets
from Bing Liu and collaborators

```{r, echo = T}
get_sentiments("bing")

```


## How does the sentiment evolves throughout the novels?
```{r, echo = T,warning=F, message=F}
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = chapter, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

## How does the sentiment evolves throughout the novels?
```{r, echo = T, eval = F,warning=F, message=F}
ggplot(jane_austen_sentiment, aes(x = index, y = sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## How does the sentiment evolves throughout the novels?
```{r, eval = T,warning=F, message=F}
ggplot(jane_austen_sentiment, aes(x = index, y = sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## WordCloud
```{r, echo = T, eval = F,warning=F, message=F}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

## WordCloud
```{r, eval = T, warning=F, message=F}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

## Most common positive and negative words
```{r, echo = T, warning=F, message=F}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

```
## Most common positive and negative words

```{r}
bing_word_counts
```
## Most common positive and negative words
```{r, echo = T, eval = F,warning=F, message=F}
bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
```

## Most common positive and negative words
```{r, echo = F, eval = T, message = F, warning=F}
bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
```

## Most common words for different emotions
```{r, echo = F, eval = T, message = F, warning = F}
nrc_word_counts <- tidy_books %>%
  inner_join(get_sentiments("nrc")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

nrc_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
```

#Thanks for your attention! Any questions?

A more detailed explanation and analysis can be found in <https://www.tidytextmining.com/>